---
title: "project1.Rmd"
author: "fatima86"
date: "January 9, 2021"
output: html_document
---
## Introduction
The data for this assignment can be downloaded from the course web site:

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) 

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as ğ™½ğ™°) </br>
date: The date on which the measurement was taken in YYYY-MM-DD format </br>
interval: Identifier for the 5-minute interval in which measurement was taken </br>
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Loading and preprocessing the data
```{r}

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
# unzip data and read 
unzip("activity.zip")
stepdata <- read.csv("activity.csv", header = TRUE)
head(stepdata)

```

## What is mean total number of steps taken per day?
```{r}
library(magrittr)
library(dplyr)

databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% summarize(tsteps= sum(steps)) %>%na.omit()
hist(databydate$tsteps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", breaks = 20)
```

## Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(databydate$tsteps)

median(databydate$tsteps)

```

## What is the average daily activity pattern?
```{r}
library(ggplot2)
databyinterval <- stepdata %>% select(interval,steps) %>% na.omit() %>% group_by(interval) %>% summarize(tsteps = mean(steps)) 
ggplot(databyinterval, aes(x=interval, y=tsteps))+ geom_line()
```
## The 5-minute interval that, on average, contains the maximum number of steps
```{r}
databyinterval[which(databyinterval$tsteps== max(databyinterval$tsteps)),]
```
## Calculate and report the total number of missing values in the dataset 
```{r}
missingVals <- sum(is.na(stepdata))
missingVals
```

## Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r}
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```
## Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)
```
## Summary of new data : mean & median

```{r}
summary(FullSummedDataByDay)
```

## Making a histogram

```{r}
hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```

## Compare the mean and median of Old and New data
```{r}
oldmean <- mean(databydate$tsteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)
# Old mean and New mean
oldmean
```

```{r}
newmean
```

```{r}
oldmedian <- median(databydate$tsteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)
# Old median and New median
oldmedian
newmean
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )
library(ggplot2)
meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")

ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval")
```